---
title: "Example Analysis"
author: "Darren Lin"
date: "2022-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
library(tidytuesdayR)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggalt)
library(purrr)
library(scales)

if(!dir.exists("../data")){
  dir.create("../data")
}

if(!file.exists("../data/coffee.RDS")) {
  tuesdata <- tt_load("2020-07-07")
  coffee <- tuesdata$coffee_ratings

  saveRDS(coffee, file = "../data/coffee.RDS")
}

coffee <- readRDS("../data/coffee.RDS")
```

# Goal

The question I'm hoping to answer is: What factors are associated with high graded coffee beans?

The data comes from TidyTuesday ([2020-07-07](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md)) The data dictionary is located [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md#coffee_ratingscsv).

# Coffee Rating and Flavor Grade

First we look at coffee ratings as it relates to the flavor grade and color of the bean.

```{r}
coffee_flv_bal <- coffee %>%
  distinct() %>%
  select(total_cup_points, flavor, balance, color) %>%
  filter(balance > 5.0) %>%
  drop_na() %>%
  group_by(flavor, color) %>%
  summarise(rating_means = mean(total_cup_points))

ggplot(coffee_flv_bal, aes(x = flavor, y = rating_means)) +
  facet_wrap(vars(color)) +
  geom_point() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(size=0.5) +
  labs(x = "Flavor Grade",
       y = "Overall Rating",
       title = "Coffee Ratings by Flavor Ratings",
       subtitle = "By Color of Bean")
  


```

While the blue-ish green and uncolored beans have a generally linear increase in bean rating as flavor grade increases, the blue-green beans seem to increase drastically in rating at around a flavor grade of 6.5-7 (out of 10), while the green beans seem to have a dip in overall rating around a flavor grade of 6 and 7.

# Coffee Rating on Certification Bodies

```{r}
coffee_cert_rate <- coffee %>%
  select(total_cup_points, certification_body) %>%
  group_by(certification_body) %>%
  summarise(rating_means = mean(total_cup_points))

coffee_cert_rate$percent <- as.double(map(coffee_cert_rate$rating_means, function(x) x/100))



ggplot(coffee_cert_rate, aes(y = reorder(certification_body, percent), x = percent)) +
  geom_lollipop(horizontal = TRUE) +
  theme(axis.text.y = element_text(size=5, angle=30)) +
  scale_y_discrete(label = abbreviate) +
  labs(x = "Percent Coffee Rating",
       y = "Cerficiation Body",
       title = "Coffee Ratings by Certification Bodies",
       subtitle = "A Lollipop Chart")
```

Next we look at coffee ratings based on certification bodies. Generally, it doesn't seem that the certification body affects a coffee bean's rating too much. METAD Agricultural Development plc seems to rate coffee beans the highest on average, and Instuto Hondereno del Cafe seems to rate coffee bean's the lowest on average (abbreviations used for the sake of space).

# Coffee Rating and Defects
```{r}
defects <- coffee %>%
  select(category_one_defects, category_two_defects) %>%
  drop_na()

defect_total <- pmap_dbl(defects, .f = sum)

defects$dfct_tot <- defect_total

coffee_dfct <- data.frame(defects, coffee$total_cup_points)

ggplot(coffee_dfct, aes(x = dfct_tot, y = coffee.total_cup_points)) + 
  geom_point() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  labs(x = "Total Defects",
       y = "Overall Rating",
       title = "Coffee Rating by Total Defects",
       subtitle = "Including Primary and Secondary Defects")
```

Finally, we look at the rating based on the number of defects. It seems the generally, the more defects a bean has the lower its rating, which is to be expected.


# Functions Used

## tidyr

`select()`

`drop_na()`

## dplyr

`distinct()`

`filter()`

`group_by()`

`summarise()`

## purr
`map()`

`pmap_dbl()`

## ggplot2

`geom_point()`

`geom_smooth()`

## ggalt
`geom_xspline()`

`geom_lollipop()`
